###testing
```{r}
library(magrittr) #for pipe operator
library(dplyr) #to use glimpse
library(ggplot2) # for graphing
library(tidyverse) #unknown
```

```{r}
chip <- read.csv('chips.csv')

chip %>% glimpse()
```
In the readme it seems there's more variables in the dictionary than just 5 so then maybe I need to read in the data manually?
```{r}
chips_manual <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-08-23/chips.csv')

chips_manual %>% 
  glimpse()
```
Didn't make a difference. Moving on. 

It may be interesting to see how the die (what even is that for CPUs and GPUs?) changes as the years go by since a major point of interest is Moors law, which is that speed/functionality of GPUs and CPUs double every year or so until reaching the limitations as specified by laws of physics.

All data types seem correct so no wrangling necessary.
```{r}
chip %>% 
  ggplot(aes(x = year, y = die_size_mm_2, color = die_size_mm_2)) +
  geom_point() +
  labs(title = "Die size of GPUs and CPUs across Time",
       x = "Year",
       y = "Die size in millimeters")
```
Seems that the size of the die increases with time.

Now how does thermal design profile relate to how many millions of transistors there are?

```{r}
chip %>% 
  ggplot(aes(x = tdp_w, y = transistors_million)) +
  geom_point() +
  labs(title = "Thermal Design Profile v Number of Millions of Transistors",
       x = "Thermal Design Profile (width?)",
       y = "Millions of Transistors")
```

Also an increasing relationship although definitely not linear. I don't know what the thermal design profile is, but I do know that GPUs have more transistors than CPUs.Although the dataset doesn't specify whether the measurements are for either of them.